---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# shsannualreport
<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
<!-- badges: end -->

:construction: :construction: :construction: 

Package is in development and includes functions, data files, and scripts for testing. When functionality is finalised, the following functions can be deleted from the R folder:

* lm_data_country_stats.R
* lm_data_key_stats.R
* lm_data_nomis_example.R
* lm_data_ons_example.R
* lm_data_region_stats.R
* lm_data_time_stats.R

The examples folder can also be deleted when other functions are finalised.

TODO As working functions are added for tables etc, they should be documented here. The process of making a function to pull and format a dataset for publication could be explained as well.

:construction: :construction: :construction:

labourmarketreports is an R package to produce markdown reports with data from the following web APIs:

* [ONS](https://developer.ons.gov.uk/)
* [NOMIS](https://www.nomisweb.co.uk/api/v01/help)

## Installation
You can install the development version of shsannualreport from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("thomascrines/labourmarketreports")
```

If you are working on SCOTS, or if the above does not work, you can install manually:

1. Go to the shsannualreport repository [master branch](https://github.com/thomascrines/labourmarketreports) on GitHub
2. Click **Clone or download** then **Download ZIP**
3. Save the file locally and unzip
4. Install with install.packages():

```{r source-installation, eval = FALSE}
install.packages("C:/DownloadDirectory/labourmarketreports-master/labourmarketreports-master", repos = NULL,
                 type="source", lib = "C:/YourLibraryPath")
```
*Library path can be seen by running `.libPaths()`*

## Usage - ONS API
If using the ONS API, so it's useful to read through the [documentation](https://developer.ons.gov.uk/) before starting to use the package. The main functions are:

* `lm_ons_request_dataset` - Requests a downloadable dataset be created by the ONS API, based on a `dataset_id`, a `dataset_edition`, and a `dataset_version`, as well as an optional `dataset_dimension` and `dataset_dimension_option` to filter on. (These can either be hard-coded or obtained using other functions described later.) A filter output ID (a unique reference to the newly created dataset) is returned by the function. This can then be passed to `lm_ons_download_dataset`.
* `lm_ons_download_dataset` - Downloads a dataset that has been created in the API using `lm_ons_request_dataset`, based on the `filter_output_id` that was returned by that function.

### Examples
To return an entire dataset from the ONS API, without filtering:
```{r, eval=FALSE}
filter_output_id <- lm_ons_request_dataset(base_uri = "https://api.beta.ons.gov.uk/v1/", 
                                           dataset_id = "labour-market", 
                                           dataset_edition = "time-series", 
                                           dataset_version = "2")

unfiltered_dataset <- lm_ons_download_dataset(base_uri = "https://api.beta.ons.gov.uk/v1/",
                                              filter_output_id = filter_output_id)
```
To return a filtered dataset from the ONS API (in this case to only return the option "16+" for the dimension "agegroups"):
```{r, eval=FALSE}
filter_output_id <- lm_ons_request_dataset(base_uri = "https://api.beta.ons.gov.uk/v1/", 
                                           dataset_id = "labour-market", 
                                           dataset_edition = "time-series", 
                                           dataset_version = "2",
                                           dataset_dimension = "agegroups",
                                           dataset_dimension_option = "16+")

filtered_dataset <- lm_ons_download_dataset(base_uri = "https://api.beta.ons.gov.uk/v1/",
                                            filter_output_id = filter_output_id)
```
### Helper functions
The above examples only work as the arguments passed in are already known. The package contains a number of helper functions to get the necessary options from the API. (The base URI used above is not taken directly from the API, but from the [documentation](https://developer.ons.gov.uk/).)

The three required arguments (`dataset_id`, `dataset_edition`, and `dataset_version`) can be obtained with the functions `lm_ons_datasets`, `lm_ons_dataset_latest_version`, and `lm_ons_dataset_editions`.

The two optional arguments used for filtering (`dataset_dimension` and `dataset_dimension_option`) can be obtained with the functions `lm_ons_dataset_dimensions` and `lm_ons_dataset_dimension_options`.

### Helper function examples

#### lm_ons_datasets
To get a dataset ID, use `lm_ons_datasets`:
```{r, eval=FALSE}
datasets <- lm_ons_datasets(base_uri = "https://api.beta.ons.gov.uk/v1/")
```
Any value in the `id` column can be used as a `dataset_id` in the other functions.
The returned dataframe can be filtered to only return desired columns with the optional `variables` argument:
```{r, eval=FALSE}
datasets <- lm_ons_datasets(base_uri = "https://api.beta.ons.gov.uk/v1/", 
                            variables = c("id", "description", "title"))
```
#### lm_ons_dataset_latest_version
To get the latest available version of a dataset, use `lm_ons_dataset_latest_version`:

```{r, eval=FALSE}
lm_ons_dataset_latest_version(base_uri = "https://api.beta.ons.gov.uk/v1/", 
                              dataset_id = "labour-market")
```
#### lm_ons_dataset_editions
To get the available editions of a dataset, use `lm_ons_editions`:

```{r, eval=FALSE}
lm_ons_dataset_editions(base_uri = "https://api.beta.ons.gov.uk/v1/", 
                        dataset_id = "labour-market")
```
#### lm_ons_dataset_dimensions (optional, for filtering only)
To get the available dimensions (or column names) of a dataset, use `lm_ons_dimensions`:

```{r, eval=FALSE}
lm_ons_dataset_dimensions(base_uri = "https://api.beta.ons.gov.uk/v1/", 
                          dataset_id = "labour-market", 
                          dataset_edition = "time-series", 
                          dataset_version = "2")
```
#### lm_ons_dataset_dimension_options (optional, for filtering only)
To get the available options of a dimension, use `lm_ons_dimension_options`:

```{r, eval=FALSE}
lm_ons_dataset_dimension_options(base_uri = "https://api.beta.ons.gov.uk/v1/", 
                          dataset_id = "labour-market", 
                          dataset_edition = "time-series", 
                          dataset_version = "2",
                          dataset_dimension = "agegroups")
```

### Full example
The following is an example of using the helper functions to get metadata of a dataset, and using this metadata to download a filtered dataset:

```{r, eval=FALSE}
base_uri <- "https://api.beta.ons.gov.uk/v1/"
datasets_variables <- c("id", "description", "title")

datasets <- lm_ons_datasets(base_uri = base_uri,
                            variables = datasets_variables)

dataset_id <- datasets$id[datasets$title == "UK Labour Market"]

dataset_edition <- lm_ons_dataset_editions(base_uri = base_uri,
                                           dataset_id = dataset_id)

dataset_version <- lm_ons_dataset_latest_version(base_uri = base_uri,
                                                 dataset_id = dataset_id)


dataset_dimensions <- lm_ons_dataset_dimensions(base_uri = base_uri,
                                                dataset_id = dataset_id,
                                                dataset_edition = dataset_edition,
                                                dataset_version = dataset_version)

example_dimension <- dataset_dimensions[3]

dataset_options <- lm_ons_dataset_dimension_options(base_uri = base_uri,
                                                    dataset_id = dataset_id,
                                                    dataset_edition = dataset_edition,
                                                    dataset_version = dataset_version,
                                                    dataset_dimension = example_dimension)

example_option <- dataset_options[1]

filter_output_id <- lm_ons_request_dataset(base_uri = base_uri,
                                           dataset_id = dataset_id,
                                           dataset_edition = dataset_edition,
                                           dataset_version = dataset_version,
                                           dataset_dimension = example_dimension,
                                           dataset_dimension_option = example_option)

filtered_dataset <- lm_ons_download_dataset(base_uri = base_uri,
                                            filter_output_id = filter_output_id)
```
## Usage - NOMIS API
Again, it's useful to read through the [NOMIS documentation](https://www.nomisweb.co.uk/api/v01/help) before starting to use the package. The main function is:

* `lm_nomis_download_dataset` - Downloads a dataset from the NOMIS API, based on a `base_uri`, and a `dataset_id`. Also accepts and optional `filter_string` and `row_limit`.

### Examples
To return an entire dataset from the NOMIS API, without filtering:
```{r, eval=FALSE}
  base_uri <- "http://www.nomisweb.co.uk/"
  dataset_id <- "NM_1_1"

  lm_nomis_download_dataset(base_uri, dataset_id)
```
To return 1000 rows of a filtered dataset from the NOMIS API (in this case to only return the option "2092957697" for the dimension "geography"):
```{r, eval=FALSE}
  base_uri <- "http://www.nomisweb.co.uk/"
  dataset_id <- "NM_1_1"
  filter_string = "?geography=2092957697"
  row_limit = 10000
  
  dataset <- lm_nomis_download_dataset(base_uri, dataset_id, filter_string = filter_string, row_limit = row_limit)
```
### Helper functions
As with the ONS examples, these examples only work if the arguments are already known. The package contains a number of helper functions to get the necessary options from the API. (The base URI used above is not taken directly from the API, but from the [NOMIS documentation](https://www.nomisweb.co.uk/api/v01/help) .)

The only required argument is `dataset_id`, which can be obtained with the function `lm_nomis_datasets`.

The `filter_string`can be created with values obtained with the functions `lm_nomis_dataset_dimensions` and `lm_nomis_dataset_dimension_options`.

### Helper function examples

#### lm_nomis_datasets
To get a dataset ID, use `lm_nomis_datasets`:
```{r, eval=FALSE}
datasets <- lm_nomis_datasets(base_uri = "http://www.nomisweb.co.uk/")
```
Returned columns can be specified with the optional `variables` argument:
```{r, eval=FALSE}
datasets <- lm_nomis_datasets(base_uri = "http://www.nomisweb.co.uk/", 
                              variables = c("id", "uri", "description.value", "name.value", "components.dimension"))
```
Any value in the `id` column can be used as a `dataset_id` in the other functions.

The returned object from `lm_nomis_datasets` is required for the `lm_nomis_dimensions` function.

#### lm_nomis_dataset_dimensions (optional, for filtering only)
To get the available dimensions (or column names) of a dataset, use `lm_nomis_dimensions`:

```{r, eval=FALSE}
datasets <- lm_nomis_datasets(base_uri = "http://www.nomisweb.co.uk/", 
                              variables = c("id", "uri", "description.value", "name.value", "components.dimension"))
dataset_id <- "NM_1_1"

dimensions <- lm_nomis_dataset_dimensions(dataset_id = dataset_id, 
                                          datasets_list = datasets)
```
#### lm_nomis_dataset_dimension_options (optional, for filtering only)
To get the available options of a dimension, use `lm_nomis_dimension_options`:

```{r, eval=FALSE}
options <- lm_nomis_dataset_dimension_options(base_uri = "http://www.nomisweb.co.uk/", 
                                              dataset_id = "NM_1_1", 
                                              dataset_dimension = "GEOGRAPHY")
```
To filter a dataset currently involves manually creating a filter string to pass to `lm_nomis_download_dataset`. In the example above, using `"geography"` as the dimension and an available option of `"2092957697"`, the filter string can be manually written: `"?geography=2092957697"`. This is a possible issue to be improved in future if necessary.
